---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# STAT 545: Homework 07
### By: Stephen Chignell

Welcome to the repository! Here you'll find all the files and associated material for [Howework 4](http://stat545.com/Classroom/assignments/hw07/hw07.html). 

For this assignment, I chose to add a function to the `foofactors` package by Jenny Bryan (original is [here](https://github.com/jennybc/foofactors)). 

I sometimes have to work with messy data where there are typos and inconsistencies as a result of data entry errors, so for HW07, I wrote a function called `fclean` which goes through all of the factor levels and trims extra white space while converting each word to title case. 

- Access my modifed (forked) version of the `foofactors` package [here](https://github.com/schignel/foofactors) 



### Overview of `foofactors`

**NOTE: This is a toy package created for expository purposes. It is not meant to actually be useful. If you want a package for factor handling, please see [forcats](https://cran.r-project.org/package=forcats).**

Factors are a very useful type of variable in R, but they can also drive you nuts. This package provides some helper functions for the care and feeding of factors.

### Installation

```{r installation, eval = FALSE}
devtools::install_github("schignel/foofactors")
```
  
### Quick demo

Binding two factors via `fbind()`:

```{r}
library(foofactors)
a <- factor(c("character", "hits", "your", "eyeballs"))
b <- factor(c("but", "integer", "where it", "counts"))
```

Simply catenating two factors leads to a result that most don't expect.

```{r}
c(a, b)
```

The `fbind()` function glues two factors together and returns factor.

```{r}
fbind(a, b)
```

Often we want a table of frequencies for the levels of a factor. The base `table()` function returns an object of class `table`, which can be inconvenient for downstream work. Processing with `as.data.frame()` can be helpful but it's a bit clunky.

```{r}
set.seed(1234)
x <- factor(sample(letters[1:5], size = 100, replace = TRUE))
table(x)
as.data.frame(table(x))
```

The `freq_out()` function returns a frequency table as a well-named `tbl_df`:

```{r}
freq_out(x)
```

Human error during data entry often leads to typos and inconsistent spacing and capitlization. This is particularly true when there are multiple individuals working on entering the same categorical data. 
```{r}
production <- factor(c("  HIGH production", "MED production  ",  "MED production  ", " low Production"))
production
```

This can be fixed by identifying individual inconsistencies after factor created, but is time consuming and cumbersome. 

The `fclean` function will go through all of the factor levels and trim extra white space and convert to title case using the `stringr` package.

```{r}
fclean(production)
```

